---
title: "Homework_1"
author: "Dayra Quinonez"
output: pdf_document
date: "2023-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

1. Create a vector of three elements (2,4,6) and name that vector `vec_a`. 
Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors 
together and name the result `vec_c`.

```{r}
vec_a <- c(2,4,6)
vec_b <- c(8,10,12)

vec_c <- vec_a + vec_b

vec_c
```


2. Create a vector, named `vec_d`, that contains only two elements (14,20). 
Add this vector to `vec_a`. What is the result and what do you think R did 
(look up the recycling rule using Google)? What is the warning message that 
R gives you?

```{r}
vec_d <- c(14,20)
vec_a < vec_d + vec_a
```

The warning message that R gives is that a longer object length is not multiple 
of shorter object length


3. Next add 5 to the vector vec_a. What is the result and what did R do? 
Why doesn't in give you a warning message similar to what you saw in the
previous problem?

```{r}
vec_a <- vec_a + 5

vec_a
```


We are adding a singular number to the vector not combining or adding two 
vectors together that may have different lengths like the last problem.


4. Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two 
different ways. 
    a) First using the `seq()` function 
    b) Using the `a:b` shortcut.
    
```{r}
seq(from=1, to=5, by=1)
```

```{r}
x <- 1:5

x
```


5. Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 
    a) Using the seq() function and 
    
```{r}
seq(from=2, to=20, by=2)
```

    b) Using the a:b shortcut and some subsequent algebra. 
    *Hint: Generate the vector 1-10 and then multiple it by 2*.
    
```{r}
x <- 1:10

x <- x * 2

x
```
6. Generate a vector of 21 elements that are evenly placed between 0 and 1 using
the `seq()` command and name this vector `x`. 

```{r}
x <- seq(from=0, to=1, by=0.05)
x

```

7. Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ 
  using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep(c(2,4,8),3)
```

8. Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
  using the `rep()` command. You might need to check the help file for rep() to 
  see all of the options that rep() will accept. In particular, look at the 
  optional argument `each=`.
  
```{r}
rep(c(2,4,8),each=4)
```


10. In this problem, we will work with the matrix 

    \[ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]\]
    
    a) Create the matrix in two ways and save the resulting matrix as `M`.
        i. Create the matrix using some combination of the `seq()` and 
        `matrix()` commands.

```{r}
x <- seq(1:15) * 2

M <- matrix(x, nrow=3, byrow=TRUE)
M
```


        ii. Create the same matrix by some combination of multiple `seq()` 
        commands and either the `rbind()` or `cbind()` command. 

```{r}
x <- seq(from=2, to=10, by=2)
y <- seq(from=12, to=20, by=2)
z <- seq(from=22, to=30, by=2)

rbind(x,y,z)

```


    b) Extract the second row out of `M`.
```{r}
rownames(M) <- c('Row1', 'Row2','Row3')

M['Row2',]


```

    c) Extract the element in the third row and second column of `M`.
```{r}
colnames(M) <- c('Col1','Col2','Col3','Col4','Col5')

M['Row3','Col2']

```

12. The following code creates a `data.frame` and then has two different methods
for removing the rows with `NA` values in the column `Grade`. Explain the 
difference between the two.

```{r}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'), Grade = c(6,8,NA,9))
 
df[ -which(  is.na(df$Grade) ), ]
df[  which( !is.na(df$Grade) ), ]
```

The outcome is the same but the difference between the two methods is the 
syntax. As for the first one the character '-' is being used to exclude the NA 
while in the second method we are utilizing a not sign '!' to exclude the NA.


## Excercises 9.5
1.Download from GitHub the data file 
[Example_5.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/
Example_5.xls). Open it in Excel and figure out which sheet of data we should 
import into R. At the same time figure out how many initial rows need to be 
skipped. Import the data set into a data frame and show the structure of the 
imported data using the `str()` command. Make sure that your data has $n=31$ 
observations and the three columns are appropriately named. If you make any 
modifications to the data file, comment on those modifications.

```{r}

example_5 <- read_excel('Example_5.xls', sheet = 'RawData', skip=4, 
                        range=("A5:C36"))

str(example_5)

```

2.Download from GitHub the data file 
[Example_3.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/
Example_3.xls). Import the data set into a data frame and show the structure of 
the imported data using the `tail()` command which shows the last few rows of a 
data table. Make sure the Tesla values are `NA` where appropriate and that both 
`-9999` and `NA` are imported as NA values. If you make any modifications to 
the data file, comment on those modifications.

```{r}
example_3 <- read_excel('Example_3.xls', sheet = 'data', range = "A1:L34", 
                        na = c('NA', -9999) )

tail(example_3)

```









